---
title: "Final Report"
subtitle: "due November 16, 2021 by 11:59 PM"
author: "Ashley Bae, Gi Chun, Jay Nolt"
date: "November 2021"
output: pdf_document
---

# Load Packages
```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(tidymodels)
library(readxl)
library(dbplyr)
library(ggplot2)
``` 

# Load Data 
```{r load-data}
load(file = "~/sta198/Jackie-Fan-Club/data/ICPSR_34363/DS0001/34363-0001-Data.rda")
#rename file
data <- da34363.0001
```
Data Citation: Battle, Juan, Pastrana, Antonio Jay, and Daniels, Jessie. Social Justice Sexuality Project: 2010 National Survey, including Puerto Rico. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2013-08-09. https://doi.org/10.3886/ICPSR34363.v1

```{r select-data, echo = FALSE}
data_filtered <- data %>%
  select(Q15B,
         Q15C,
         Q17A,
         Q17B,
         Q18A1:Q18A5,
         Q18C,
         Q18G,
         Q19A1:Q19A7,
         Q22A,
         Q22B,
         Q25)
# remove descriptions and other attributes
data_filtered2 <- lapply(data, function(x) {attributes(x) <- NULL; x}) %>% 
  as.data.frame() %>%
  select(Q15B,
         Q15C,
         Q17A,
         Q17B,
         Q18A1:Q18A5,
         Q18C,
         Q18G,
         Q19A1:Q19A7,
         Q22A,
         Q22B,
         Q25)
```

```{r renameQuestions, echo = FALSE}
data_big <- data_filtered %>%
    select(Q17A,
         Q17B,
         Q18A1:Q18A5,
         Q18C,
         Q18G,
         Q19A1:Q19A7,
         Q22A,
         Q22B,
         Q25) %>%
  rename(healthInsureAcc = Q17A) %>%
  rename(healthProvideAcc = Q17B) %>%
  rename(male = Q18A1) %>%
  rename(female = Q18A2) %>% 
  rename(m2f = Q18A3) %>%
  rename(f2m = Q18A4) %>%
  rename(genderOther = Q18A5) %>%
  rename(sexuality = Q18C) %>%
  mutate(age = 2021 - Q18G) %>%
  rename(black = Q19A1) %>%
  rename(hispanic = Q19A2) %>% 
  rename(asian = Q19A3) %>%
  rename(native = Q19A4) %>%
  rename(white = Q19A5) %>%
  rename(multi = Q19A6) %>%
  rename(raceOther = Q19A7) %>%
  rename(edu = Q22A) %>%
  rename(income = Q22B) %>%
  rename(assessHealth = Q25)
```

```{r creatingGender, echo = FALSE}
data <- data_big %>%
  mutate(male = case_when(
  male == "(1) Yes" ~ "male",
  is.na(male) ~ "",
  TRUE ~ ""), 
  female = case_when(
  female == "(1) Yes" ~ "female",
  is.na(female) ~ "",
  TRUE ~ ""), 
  m2f = case_when(
  m2f == "(1) Yes" ~ "m2f",
  is.na(m2f) ~ "",
  TRUE ~ ""), 
  f2m = case_when(
  f2m == "(1) Yes" ~ "f2m",
  is.na(f2m) ~ "",
  TRUE ~ ""), 
  genderOther = case_when(
  genderOther == "(1) Yes" ~ "Other",
  is.na(genderOther) ~ "",
  TRUE ~ ""), 
  gender = ""
) 
data$gender <- paste(data$male, data$female)
data$gender <- paste(data$gender, data$m2f)
data$gender <- paste(data$gender, data$f2m)
data$gender <- paste(data$gender, data$genderOther)
data$gender <- trimws(data$gender)
data <- data %>%
mutate(gender = case_when(
  gender == "male" ~ "Male",
  gender == "female" ~ "Female",
  gender == "m2f" ~ "Transgender: Male to Female",
  gender == "f2m" ~ "Transgender: Female to Male",
  TRUE ~ "Other")
)
```

```{r creatingRace, echo = FALSE}
data <- data %>%
  mutate(black = case_when(
  black == "(1) Yes" ~ "black",
  is.na(male) ~ "",
  TRUE ~ ""), 
  hispanic = case_when(
  hispanic == "(1) Yes" ~ "hispanic",
  is.na(hispanic) ~ "",
  TRUE ~ ""), 
  asian = case_when(
  asian == "(1) Yes" ~ "asian",
  is.na(asian) ~ "",
  TRUE ~ ""), 
  native = case_when(
  native == "(1) Yes" ~ "native",
  is.na(native) ~ "",
  TRUE ~ ""), 
  white = case_when(
  white == "(1) Yes" ~ "white",
  is.na(white) ~ "",
  TRUE ~ ""), 
  multi = case_when(
  multi == "(1) Yes" ~ "multi",
  is.na(multi) ~ "",
  TRUE ~ ""), 
  raceOther = case_when(
  raceOther == "(1) Yes" ~ "Other",
  is.na(raceOther) ~ "",
  TRUE ~ ""), 
  race = ""
) 
data$race <- paste(data$black, data$hispanic)
data$race <- paste(data$race, data$asian)
data$race <- paste(data$race, data$native)
data$race <- paste(data$race, data$white)
data$race <- paste(data$race, data$multi)
data$race <- paste(data$race, data$raceOther)
data$race <- trimws(data$race)
data <- data %>%
mutate(race = case_when(
  race == "black" ~ "Black",
  race == "hispanic" ~ "Hispanic",
  race == "asian" ~ "Asian",
  race == "native" ~ "Native",
  race == "white" ~ "White",
  race == "multi" ~ "Multiracial",
  race == "other" ~ "Other",
  race == "" ~ "Other",
  TRUE ~ "Multiracial")
)
```


# Data Analysis 

To evaluate statistically significant impacts on health outcomes due to sexuality, gender, race, and intersectional identities, we ran a few statistical tests. 

```{r preliminary-regression, echo = FALSE, message = FALSE, warning = FALSE} 
data_fit <- data %>% 
mutate(sexuality_queer = if_else(sexuality == "(08) Straight/Heterosexual", "0", "1"),
       gender_trans = if_else((gender == "Male") | (gender == "Female"), "0", "1"), 
       WHITE = if_else(white == "(1) Yes", 1, 0),
       BLACK = if_else(black == "(1) Yes", 1, 0),
       HISPANIC = if_else(hispanic == "(1) Yes", 1, 0),
       ASIAN = if_else(asian == "(1) Yes", 1, 0),
       NATIVE = if_else(native == "(1) Yes", 1, 0),
       MULTIRACE = if_else(multi == "(1) Yes", 1, 0), 
       RACEOTHER = if_else(raceOther == "(1) Yes", 1, 0)) 

data_relevel <- data_fit
data_relevel$gender_trans = relevel(as.factor(data_relevel$gender_trans), ref = "0") 
data_relevel$race = relevel(as.factor(data_relevel$race), ref = "White") 
data_relevel$sexuality_queer = relevel(as.factor(data_relevel$sexuality_queer), ref = "0") 

# insurance fit ---------------------------------------------------------------------
insurancefit <- glm(healthInsureAcc ~ race + gender_trans + sexuality_queer,
           data = data_relevel, family = "binomial")
summary(insurancefit, conf.int = TRUE, exponentiate = TRUE)
# access fit ---------------------------------------------------------------------
providerfit <- glm(healthProvideAcc ~ race + gender_trans + sexuality_queer,
           data = data_relevel, family = "binomial")
summary(providerfit, conf.int = TRUE, exponentiate = TRUE)


# insurance fit interactions  ---------------------------------------------------------------------
insurancefit_interact <- glm(healthInsureAcc ~ race*gender_trans + gender_trans*sexuality_queer + sexuality_queer*race,
           data = data_relevel, family = "binomial")
summary(insurancefit_interact, conf.int = TRUE, exponentiate = TRUE)

# access fit interactions  ---------------------------------------------------------------------
providerfit_interact <- glm(healthProvideAcc ~ race*gender_trans + gender_trans*sexuality_queer + sexuality_queer*race,
           data = data_relevel, family = "binomial")
summary(providerfit_interact, conf.int = TRUE, exponentiate = TRUE)

# ANOVA for better fit -------------------------------------------
insurance_anova <- anova(insurancefit, insurancefit_interact, test = "LRT") 
summary(insurance_anova)

provider_anova <- anova(providerfit, providerfit_interact, test = "LRT") 
summary(provider_anova)
```
We ran a multiple regression for each health insurance access and health provider access conditional on gender, race, and sexuality. We saw that our referent group of white, cis, heterosexual individuals are 1.47 as likely to have health insurance than non-white, trans, non-heterosexual counterparts and 1.32 times as likely for access to a regular healthcare provider.

Sexuality is a complicated variable to work with, given that the question posed in the survey offers 12 options that are not all mutually exclusive, yet the answer given must be only one of the above. Also, to retain all options given would overcomplicate and weaken the model. Therefore, we decided to distinguish between heterosexual and all other sexuality answers in our subsequent models. 

In healthcare access, we found that the most negatively affected race group was Hispanic and the most negatively affected gender group was transgender individuals.  

We conducted another multiple regression with interaction terms, in order to evaluate intersection of gender and race groups. 

```{r overall-anova}
summary(aov(Q18C~Q25,data=data_filtered2))
```
```{r stepdown-anova}
sexpair <- pairwise.t.test(data_filtered2$Q18C,data_filtered2$Q25, p.adj = "holm")
sigpairs <- broom::tidy(sexpair) %>%
  filter(p.value<0.05) %>%
  arrange(group1,group2)
nrow(sigpairs)
print(sigpairs) 
```
4: 2-spirit
5: queer

